Reanalysis Step 6: Alternative Propensity Score Specifications

### 2025-08-27 reanalysis-4-outcome_analysis.qmd
wip

@@

  # Calculate AME
  ame <- avg_slopes(
    fitted_models[[i]], 
    variables = "ever_lapse_binary",
    vcov = "HC3",
    wts = "weights"
  )

@@

# Pool AME estimates
pooled_ame <- data.frame(
  Effect = "Average Marginal Effect of Lapse",
  AME = mean(ame_combined$ame),
  SE = sqrt(mean(ame_combined$se^2) + var(ame_combined$ame) * (1 + 1/n_ame_imps)),
  Interpretation = paste0(
    "On average, lapsing from care increases the probability of VI/blindness by ",
    round(mean(ame_combined$ame) * 100, 1), " percentage points"
  )
)

kable(pooled_ame |> select(-Interpretation), 
      caption = "Average marginal effect of lapsing")

cat("\n", pooled_ame$Interpretation, "\n")
```

@@

message("Results saved to: ", 
        file.path(reanalysis_data_dir, "outcome_analysis_main_effects_results.rds"))
```

@@

- **Primary finding**: Lapsing from care has a significant effect on vision impairment/blindness (OR = `r round(lapse_effect$OR, 2)`, p `r ifelse(lapse_effect$p.value < 0.001, "< 0.001", paste("=", round(lapse_effect$p.value, 3)))`)
- **Effect size**: The average marginal effect indicates that lapsing increases the probability of VI/blindness by approximately `r round(mean(ame_combined$ame) * 100, 1)` percentage points
- **Consistency**: The effect is present across all levels of DR severity and treatment status
- **Key insight**: Even patients with no DR and no treatment show increased risk when lapsing from care

>>>>

# Check for required files
required_files <- c(
  "imputed_datasets.rds",
  "twang_weights.rds",
  "variable_lists.rds"
)

missing_files <- c()
for (file in required_files) {
  if (!file.exists(file.path(reanalysis_data_dir, file))) {
    missing_files <- c(missing_files, file)
  }
}

if (length(missing_files) > 0) {
  stop(paste("Required files not found. Please run previous steps first:\n",
             "- Step 2 (Multiple Imputation) creates: imputed_datasets.rds\n",
             "- Step 3 (Propensity Scores) creates: twang_weights.rds\n",
             "Missing files:", paste(missing_files, collapse = ", ")))
}

@@

  # Create survey design to properly handle clustering
  design <- svydesign(
    ids = ~ e_mrn_deidentified,  # Cluster on patient ID
    weights = ~ weights,
    data = imp_data
  )

  # Calculate AME with survey design
  # Note: marginaleffects handles survey objects appropriately
  ame <- avg_slopes(
    fitted_models[[i]], 
    variables = "ever_lapse_binary",
    newdata = imp_data,  # Use the full dataset
    vcov = ~e_mrn_deidentified,  # Cluster standard errors on patient ID
    wts = imp_data$weights  # Apply propensity score weights
  )

@@

# Pool AME estimates using Rubin's rules
pooled_ame <- data.frame(
  Effect = "Average Marginal Effect of Lapse",
  AME = mean(ame_combined$ame),
  SE = sqrt(mean(ame_combined$se^2) + var(ame_combined$ame) * (1 + 1/n_ame_imps)),
  Interpretation = paste0(
    "On average, lapsing from care increases the probability of VI/blindness by ",
    round(mean(ame_combined$ame), 3)
  )
)

kable(pooled_ame |> select(-Interpretation), 
      caption = "Average marginal effect of lapsing",
      digits = 4)

# Display interpretation separately
ame_text <- pooled_ame$Interpretation
```

`r ame_text`

@@

save_location <- file.path(reanalysis_data_dir, "outcome_analysis_main_effects_results.rds")
```

**Results saved to:** `r save_location`

@@

- **Primary finding**: Lapsing from care has a significant effect on vision impairment/blindness (OR = `r round(lapse_effect$OR, 2)`, p `r ifelse(lapse_effect$p.value < 0.001, "< 0.001", paste("=", round(lapse_effect$p.value, 3)))`)
- **Effect size**: The average marginal effect indicates that lapsing increases the probability of VI/blindness by `r round(mean(ame_combined$ame), 3)`
- **Consistency**: The effect is present across all levels of DR severity and treatment status
- **Key insight**: Even patients with no DR and no treatment show increased risk when lapsing from care

### 2025-08-27 reanalysis-5-sensitivity.qmd
wip

>>>>

# Check for required files
required_files <- c(
  "outcome_analysis_main_effects_results.rds",
  "imputed_datasets.rds",
  "twang_weights.rds"
)

missing_files <- c()
for (file in required_files) {
  if (!file.exists(file.path(reanalysis_data_dir, file))) {
    missing_files <- c(missing_files, file)
  }
}

if (length(missing_files) > 0) {
  stop(paste("Required files not found. Please run previous steps first:\n",
             "- Step 2 (Multiple Imputation) creates: imputed_datasets.rds\n",
             "- Step 3 (Propensity Scores) creates: twang_weights.rds\n",
             "- Step 4 (Outcome Analysis) creates: outcome_analysis_main_effects_results.rds\n",
             "Missing files:", paste(missing_files, collapse = ", ")))
}

### 2025-08-27 reanalysis-6-alternative_specifications.qmd
wip

>>>>

# Check for required files
required_files <- c(
  "imputed_datasets.rds",
  "variable_lists.rds"
)

missing_files <- c()
for (file in required_files) {
  if (!file.exists(file.path(reanalysis_data_dir, file))) {
    missing_files <- c(missing_files, file)
  }
}

if (length(missing_files) > 0) {
  stop(paste("Required files not found. Please run previous steps first:\n",
             "- Step 1 (Data Preparation) creates: variable_lists.rds\n",
             "- Step 2 (Multiple Imputation) creates: imputed_datasets.rds\n",
             "Missing files:", paste(missing_files, collapse = ", ")))
}
